---
title: "CSLS Lake Evaporation Method Comparison"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{csls_lake_evap_method_comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r calculate_evap, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(CSLSevap)
library(lubridate)
library(dplyr)
library(reshape2)
library(ggplot2)

lakes    <- c("Pleasant", "Long", "Plainfield")
methods  <- c("McJannet", "McJannet", "Hamon")
use_lsts <- c(TRUE, FALSE, FALSE)
names    <- c("McJannet_LST", "McJannet_no_LST", "Hamon")

all_E  <- NULL
for (i in 1:length(methods)) {
  method    <- methods[i]
  use_lst   <- use_lsts[i]
  daily_met <- CSLS_daily_met(method = method,
                              use_lst = use_lst)
  monthly_E <- daily_met %>%
               group_by(lake = .data$lake,
                        date = floor_date(.data$date, unit = "month")) %>%
               summarise(E_mm = sum(.data$E)) %>%
               ungroup() %>%
               select(.data$lake, .data$date, .data$E_mm)
  if (is.null(all_E)) {
    all_E <- monthly_E
  } else {
    all_E <- merge(all_E, monthly_E, by = c("lake", "date"))
  }
}
colnames(all_E) <- c("lake", "date", names)
all_E     <- all_E %>% filter(date != mdy("06-01-2018"))
melted_E  <- melt(all_E, id.vars = c("date", "lake"))
```

## Overview

This document compares possible options for calculating lake evaporation for the Central Sands Lakes Study. These options include:

* **McJannet_LST:** Lake evaporation based on McJannet et al. (2008) with daily lake surface temperatures incorporated. Only feasible for 2018-2019.
* **McJannet_no_LST:** Lake evaporation based on McJannet et al. (2008) with daily lake surface temperatures estimated.
* **Hamon:** Lake evaporation using the Unmodified Hamon method (Harwell, 2012), which has fewer required inputs.

<br>

Visually, these methods capture similar overall trends in monthly lake evaporation. It appears that the Hamon method probably works best for lakes that are smaller than the CSLS lakes. The larger a lake is, the slower it is to respond to changes in air temperature (warming up or cooling down). Thus, the Hamon estimates are a closer match to the McJannet estimates for Long Lake and Plainfield Lake (the smaller CSLS lakes), and not as good of a match for Pleasant Lake.

```{r plot, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 7.5, fig.height = 4.5}
ggplot(data = melted_E, 
       aes(x = date, y = value, color = variable)) +
  geom_line() + 
  facet_wrap(~lake) +
  scale_x_datetime(breaks = "4 months",
                   date_labels = "%b '%y") +
  scale_color_manual(values = c("black", "grey70", "red")) +
  labs(x = "", y = "Evaporation (mm)", color = "Method") +
  theme_bw() +
  theme()
```
